Income and Food Price Transmission

### 2016-08-11 appendix.Rnw
JAE sub. vers.

{\small{}}%
\begin{tabular}{|c|>{\raggedright}m{8cm}|>{\raggedright}m{3cm}|}
\hline 
\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} \tabularnewline
\hline
{\small{}FPI} & {\small{}Domestic food price index.} & {\small{}ILO, LABORSTA} \tabularnewline
\hline 
{\small{}RFPI} & {\small{}Domestic real food price index. FPI deflated by the CPI} & {\small{}ILO, LABORSTA, own calculations} \tabularnewline
\hline 
{\small{}FAO FPI} & {\small{}FAO food price index.} & {\small{}FAO} \tabularnewline
\hline 
{\small{}FAO RFPI} & {\small{}Real FAO food price index. Nominal index deflated by the
World Bank Manufactures Unit Value index} & {\small{}FAO} \tabularnewline
\hline 
{\small{}ER} & {\small{}Exchange rate. Log of the domestic currency/US\$ exchange rate} & {\small{}IMF} \tabularnewline
\hline 
{\small{}GDP} & {\small{}GDP. Log of per capita GDP (constant 2011 international \$, PPP)} & {\small{}World Bank database} \tabularnewline
\hline 
{\small{}CCONS} & {\small{}Cereal consumption share. Share of dietary energy supply derived from
cereals, roots and tubers (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}CIMP} & {\small{}Cereal import share. Cereal imports as a share of domestic supply (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}CEXP} & {\small{}Cereal export share. Cereal exports as a share of domestic production (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}LLDC} & {\small{}Dummy. 1 if country is a landlocked developing country} & {\small{}UNCTAD} \tabularnewline
\hline
{\small{}STE} & {\small{}State trading enterprise dummy. 1 if country has a grain state trading enterprise} & {\small{}\citet{greb12}} \tabularnewline
\hline
{\small{}TO} & {\small{}Trade openness. Sum of exports and imports of goods and services as a share of GDP (\%). 2006-10 average} & {\small{}World Bank Development Indicators} \tabularnewline
\hline 
{\small{}DTF} & {\small{}Distance to Frontier index. An economy\textquoteright s distance to frontier measured on a scale from 0 to 100, where 0 represents the lowest performance and 100 the frontier.} & {\small{}Doing Business survey, Trading Across Borders. World Bank} \tabularnewline
\hline 
{\small{}LPI} & {\small{} Logistics performance index: Quality of trade and transport-related infrastructure. The index ranges from 1 to 5, with a higher score representing better performance.} & {\small{}Logistics Performance Index 2007. World Bank database} \tabularnewline
\hline 
\end{tabular}{\small \par}
\end{center}
\end {table}

<<results=tex,echo=F>>=
library(stargazer)
reg.list = readRDS("int_reg_m.rds")
se.list = readRDS("int_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2007-8', 
          label = "tab:int-regressions",
          dep.var.labels=c("FPI inflation", "RFPI growth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$",
                             "$Export$",
                             "$Price$",
                             "$Stock$",
                             "$Tariff$",
                             "$Tax$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}. Footnote \\ref{int_note} contains a description", 
                  "of the 5 intervention dummies."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
# library(stargazer)
reg.list = readRDS("fc_reg1011_m.rds")
se.list = readRDS("fc_reg1011_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2010-11', 
          label = "tab:1011-regressions",
          dep.var.labels=c("FPI inflation", "RFPI grwth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
         covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS("rfpi_pt_reg_m.rds")
se.list = readRDS("rfpi_pt_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run real FPI transmission rates', 
          label = "tab:rfpi-pt-regressions", 
          dep.var.labels=c("RFPI LRM (6 lags)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS("fpi_pt_reg_m1.rds")
se.list = readRDS("fpi_pt_reg_se1.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run FPI transmission rates', 
          label = "tab:fpi-pt1-regressions", 
          dep.var.labels=c("FPI LRM (single lag)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS("rfpi_pt_reg_m1.rds")
se.list = readRDS("rfpi_pt_reg_se1.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run real FPI transmission rates', 
          label = "tab:rfpi-pt1-regressions", 
          dep.var.labels=c("RFPI LRM (single lag)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
         notes=c("*, ** and *** signify significance at the 10, 5 and 1 percent",
                  "levels, respectively. Robust standard errors in parentheses;", 
                  "samples exclude outliers defined as observations $\\pm$ 3.5 stan-", 
                  "dard deviations from the mean of the dependent variable.", 
                  "A description of the variables and their sources can be found", 
                  "in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

>>>>

{\small{}}%
\begin{tabular}{|c|>{\raggedright}m{8cm}|>{\raggedright}m{3cm}|}
\hline 
\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} \tabularnewline
\hline
{\small{}FPI} & {\small{}Domestic food price index.} & {\small{}ILO, LABORSTA} \tabularnewline
\hline 
{\small{}RFPI} & {\small{}Domestic real food price index. FPI deflated by the CPI} & {\small{}ILO, LABORSTA, own calculations} \tabularnewline
\hline 
{\small{}FAO FPI} & {\small{}FAO food price index.} & {\small{}FAO} \tabularnewline
\hline 
{\small{}FAO RFPI} & {\small{}Real FAO food price index. Nominal index deflated by the
World Bank Manufactures Unit Value index} & {\small{}FAO} \tabularnewline
\hline 
{\small{}ER} & {\small{}Exchange rate. Log of the domestic currency/US\$ exchange rate} & {\small{}IMF} \tabularnewline
\hline 
{\small{}GDP} & {\small{}GDP. Log of per capita GDP (constant 2011 international \$, PPP)} & {\small{}World Bank database} \tabularnewline
\hline 
{\small{}CCONS} & {\small{}Cereal consumption share. Share of dietary energy supply derived from
cereals, roots and tubers (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}CIMP} & {\small{}Cereal import share. Cereal imports as a share of domestic supply (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}CEXP} & {\small{}Cereal export share. Cereal exports as a share of domestic production (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} \tabularnewline
\hline 
{\small{}LLDC} & {\small{}Dummy. 1 if country is a landlocked developing country} & {\small{}UNCTAD} \tabularnewline
\hline
{\small{}STE} & {\small{}State trading enterprise dummy. 1 if country has a grain state trading enterprise} & {\small{}\citet{Greb2016}} \tabularnewline
\hline
{\small{}TO} & {\small{}Trade openness. Sum of exports and imports of goods and services as a share of GDP (\%). 2006-10 average} & {\small{}World Bank Development Indicators} \tabularnewline
\hline 
{\small{}DTF} & {\small{}Distance to Frontier index. An economy\textquoteright s distance to frontier measured on a scale from 0 to 100, where 0 represents the lowest performance and 100 the frontier.} & {\small{}Doing Business survey, Trading Across Borders. World Bank} \tabularnewline
\hline 
{\small{}LPI} & {\small{} Logistics performance index: Quality of trade and transport-related infrastructure. The index ranges from 1 to 5, with a higher score representing better performance.} & {\small{}Logistics Performance Index 2007. World Bank database} \tabularnewline
\hline 
\end{tabular}{\small \par}
\end{center}
\end {table}

<<results=tex,echo=F>>=
library(stargazer)
str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
reg.list = readRDS(paste0(str,"int_reg_m.rds"))
se.list = readRDS(paste0(str,"int_reg_se.rds"))

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2007:1-2008:6', 
          label = "tab:int-regressions",
          dep.var.labels=c("FPI inflation", "RFPI growth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$",
                             "$Export$",
                             "$Price$",
                             "$Stock$",
                             "$Tariff$",
                             "$Tax$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
                  "percent levels, respectively. Robust standard errors in;", 
                  "parentheses samples exclude outliers defined as obser-", 
                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
                  "dependent variable. A description of the variables and", 
                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
# library(stargazer)
reg.list = readRDS(paste0(str,"fc_reg1011_m.rds"))
se.list = readRDS(paste0(str,"fc_reg1011_se.rds"))

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2010:1-2010:12', 
          label = "tab:1011-regressions",
          dep.var.labels=c("FPI inflation", "RFPI growth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
         covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
                  "percent levels, respectively. Robust standard errors in;", 
                  "parentheses samples exclude outliers defined as obser-", 
                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
                  "dependent variable. A description of the variables and", 
                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
reg.list = readRDS(paste0(str,"rfpi_pt_reg_m.rds"))
se.list = readRDS(paste0(str,"rfpi_pt_reg_se.rds"))

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run real FPI transmission rates', 
          label = "tab:rfpi-pt-regressions", 
          dep.var.labels=c("RFPI LRM (6 lags)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
                  "percent levels, respectively. Robust standard errors in;", 
                  "parentheses samples exclude outliers defined as obser-", 
                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
                  "dependent variable. A description of the variables and", 
                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

### 2016-08-11 conclusion.Rnw
JAE sub. vers.

In the third and final approach the determinants of short run price transmission are estimated directly with a dynamic panel model. Results indicate that short run food price transmission is affected by the many of the same factors as as long run transmission and in the same manner. I conclude that countries from the middle of the income spectrum generally have the highest food price transmission rates once we control for factors that are commonly believed to affect price transmission. This is a worrying finding in a way, given the fact that nearly three quarters of the world's poor now live in middle-income countries \citep{worldbank16, Sumner2012, Kanbur2012}.

>>>>

In the third and final approach the determinants of short run price transmission are estimated directly with a dynamic panel model. Results indicate that short run food price transmission is affected by the many of the same factors as as long run transmission and in the same manner. I conclude that countries from the middle of the income spectrum generally have the highest food price transmission rates once we control for factors that are commonly believed to affect price transmission. This is a worrying finding in a way, given the fact that nearly three quarters of the world's poor now live in middle-income countries \citep{Worldbank2016, Sumner2012, Kanbur2012}.
